---
title: "03 DEG Analysis - Combined Samples"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "03_DEG_comb.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-multiome-visium](https://github.com/eturkes/tau-mutant-multiome-visium).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-multiome-visium.
#    Copyright (C) 2023  Emir Turkes, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(DT::JS, .quiet = TRUE)
packages <- c(
  "Seurat", "SingleCellExperiment", "scuttle", "parallelly", "ggplot2", "dplyr", "ggrepel", "viridis", "DT", "patchwork"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
analysis_no <- 3
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset derived from path.

cache_dir <- file.path("..", "..", "tmp", "cache", data_name, paste0("0", analysis_no)) # Intermediate data objects.
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "02", "seurat.rds"))
seurat
```

# NLF S305N+3 vs. S305N+3

```{r}
ident1 <- "NLF_S305N+3"
ident2 <- "S305N+3"

seurat@active.ident <- factor(seurat$group)
markers <- FindMarkers(seurat, ident.1 = ident1, ident.2 = ident2, verbose = FALSE)
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(ident1, ident2))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample")

seurat$seurat_clusters_group <- factor(paste(seurat$seurat_clusters, seurat$group, sep = "_"))
seurat@active.ident <- seurat$seurat_clusters_group
```

## Oligodendrocytes

```{r}
cluster_num <- 0

markers <- FindMarkers(
  seurat, ident.1 = paste(cluster_num, ident1, sep = "_"),
  ident.2 = paste(cluster_num, ident2, sep = "_"), verbose = FALSE
)
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(paste(cluster_num, ident1, sep = "_"), paste(cluster_num, ident2, sep = "_")))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample") + plot_annotation("Oligodendrocytes")
```

## Microglia

*NOTE: No multiple testing correction*

```{r}
cluster_num <- 9

markers <- FindMarkers(
  seurat, ident.1 = paste(cluster_num, ident1, sep = "_"),
  ident.2 = paste(cluster_num, ident2, sep = "_"), verbose = FALSE
)
markers <- markers[markers$p_val < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(paste(cluster_num, ident1, sep = "_"), paste(cluster_num, ident2, sep = "_")))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample") + plot_annotation("Microglia")
```

## Astrocytes

```{r}
cluster_num <- 2

markers <- FindMarkers(
  seurat, ident.1 = paste(cluster_num, ident1, sep = "_"),
  ident.2 = paste(cluster_num, ident2, sep = "_"), verbose = FALSE
)
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(paste(cluster_num, ident1, sep = "_"), paste(cluster_num, ident2, sep = "_")))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample") + plot_annotation("Astrocytes")
```

## Cortical Glutamatergic

```{r}
cluster_num <- 1

markers <- FindMarkers(
  seurat, ident.1 = paste(cluster_num, ident1, sep = "_"),
  ident.2 = paste(cluster_num, ident2, sep = "_"), verbose = FALSE
)
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(paste(cluster_num, ident1, sep = "_"), paste(cluster_num, ident2, sep = "_")))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample") + plot_annotation("Ctx Glutamatergic")
```

## GABAergic

```{r}
cluster_num <- 8

markers <- FindMarkers(
  seurat, ident.1 = paste(cluster_num, ident1, sep = "_"),
  ident.2 = paste(cluster_num, ident2, sep = "_"), verbose = FALSE
)
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download_exp(markers)

seurat_sub <- subset(seurat, idents = c(paste(cluster_num, ident1, sep = "_"), paste(cluster_num, ident2, sep = "_")))
VlnPlot(seurat_sub, features = rownames(markers)[1:6], group.by = "sample") + plot_annotation("GABAergic")
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
